---
title: "README"
output: github_document
always_allow_html: true

#Bhan, Lam (2022) [Nanyang Technological University]
#ORCID: 0000-0001-5193-6560
#GitHub: https://github.com/bhanlam/
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(pander)
library(dataverse)
library(dplyr)
library(stringr)
library(kableExtra)
library(tidyr)
library(rstatix)

source("helper.R")
```

This repository contains the code accompanying the publication "Crossing the Linguistic Causeway: A Binational Approach for Translating Soundscape Attributes to Bahasa Melayu". This paper is hosted on [arXiv]() and the raw dataset is hosted at [10.21979/N9/0NE37R](https://doi.org/10.21979/N9/0NE37R).

## Environment

The code has been tested on the following platform.

```{r session, message=FALSE, echo=FALSE}
pander(sessionInfo())
```

## Data preparation

The raw data file is retrieved and cleaned.

### Dataloader
The raw dataset is retrieved from the DR-NTU (Data) [https://researchdata.ntu.edu.sg] Dataverse repository at [10.21979/N9/0NE37R](https://doi.org/10.21979/N9/0NE37R) and processed.

```{r dataloader, message=FALSE}
#Set dataverse server
Sys.setenv("DATAVERSE_SERVER" = "https://researchdata.ntu.edu.sg")
#Retrieve raw data in csv
dataset <- get_dataframe_by_name(
  filename = "SATPQn.csv", 
  dataset = "doi:10.21979/N9/0NE37R",
  original = TRUE,
  .f = read.csv)
```
### Data cleaning

The raw data headers are first converted to ASCII to remove UTF-8 formatting. The survey responses are subsetted to remove participant data with "country of residence" that is neither Singapore (SG) or Malaysia (MY). The resulting dataset consists of 63 participant responses after the removal of 3 participants. 

This is followed by extraction of individual dataframes corresponding to the respective criteria categories for main and derived axis attributes, respectively. 

```{r cleaning, message=FALSE}
#convert column names to ASCII
colnames(dataset)<-iconv(colnames(dataset),
                        from = 'UTF-8', 
                        to = 'ASCII//TRANSLIT') 

#remove non-SG/MY respondents
sgmydata <- dataset %>%
        dplyr::rename(COUNTRY=Current.country.of.residence) %>%
        mutate(COUNTRY=ifelse(COUNTRY=="Malaysia","MY", COUNTRY)) %>%
        mutate(COUNTRY=ifelse(COUNTRY=="Singapore","SG", COUNTRY)) %>%
        mutate(COUNTRY=ifelse(COUNTRY=="Europe","Others", COUNTRY)) %>%
        mutate(COUNTRY=ifelse(COUNTRY=="Asia (excl. MY and SG)",
                              "Others", COUNTRY)) %>%
        filter(!str_detect(COUNTRY, "Others"))

#define column names for main and derived axes
mainColnames<-c('COUNTRY','APPR','UNDR','ANTO','BIAS',
                'ASSOCCW','IMPCCW','ASSOCW','IMPCW','CANDIDATE')

derivedColnames<-c('COUNTRY','APPR','UNDR','ASSOCCW',
                   'IMPCCW','ASSOCW','IMPCW','CANDIDATE')

#extract responses for each PAQ attribute and rename columns with criteria
#main axes
eventdf<-sgmydata %>% 
        select(matches('that.meriah|is.meriah|COUNTRY')) %>%
        mutate(candidate="meriah") %>% setNames(mainColnames)

pleasdf<-sgmydata %>%
        select(matches('menyenangkan|COUNTRY')) %>%
        mutate(candidate="menyenangkan") %>% setNames(mainColnames)

uneventdf<-sgmydata %>%
        select(matches('tidak.meriah|COUNTRY')) %>%
        mutate(candidate="tidak meriah") %>% setNames(mainColnames)

annoydf<-rbind(sgmydata %>%
                select(matches('membingitkan|COUNTRY')) %>%
                mutate(candidate="membingitkan") %>% 
                setNames(mainColnames),
               sgmydata %>%
                select(matches('menjengkelkan|COUNTRY')) %>%
                mutate(candidate="menjengkelkan") %>% 
                setNames(mainColnames))

#dervied axes
vibrantdf<-rbind(sgmydata %>% select(matches('rancak|COUNTRY')) %>%
                         mutate(candidate="rancak") %>% 
                         setNames(derivedColnames),
                 sgmydata %>% select(matches('bersemarak|COUNTRY')) %>%
                         mutate(candidate="bersemarak") %>% 
                         setNames(derivedColnames))

calmdf<-rbind(sgmydata %>% select(matches('tenang|COUNTRY')) %>%
                         mutate(candidate="tenang") %>% 
                         setNames(derivedColnames),
                 sgmydata %>% select(matches('menenangkan|COUNTRY')) %>%
                         mutate(candidate="menenangkan") %>% 
                         setNames(derivedColnames))

monotdf<-rbind(sgmydata %>% select(matches('that.membosankan|COUNTRY')) %>%
                         mutate(candidate="membosankan") %>% 
                         setNames(derivedColnames),
                sgmydata %>% select(matches('tidak.berubah.oleh.itu.membosankan|COUNTRY')) %>%
                         mutate(candidate="tidak berubah oleh itu membosankan") %>% 
                         setNames(derivedColnames),
               sgmydata %>% select(matches('kurang.kepelbagaian.oleh.itu.membosankan|COUNTRY')) %>%
                         mutate(candidate="kurang kepelbagaian oleh itu membosankan") %>% 
                         setNames(derivedColnames))

chaoticdf<-rbind(sgmydata %>% select(matches('huru.hara|COUNTRY')) %>%
                         mutate(candidate="huru-hara") %>% 
                         setNames(derivedColnames),
                 sgmydata %>% select(matches('kelam.kabut|COUNTRY')) %>%
                         mutate(candidate="kelam-kabut") %>% 
                         setNames(derivedColnames))
```

### Cleaning of demographic information

```{r democlean, message=FALSE}
#column names
demoCol <- c("Prof.zsm","Prof.eng","COUNTRY","LoSOut","DISCIPLINE")

sgmydata_demo <- sgmydata %>%
        select(2:6) %>% setNames(demoCol) %>%
        mutate(DISCIPLINE=ifelse(grepl('*Science*',DISCIPLINE, ignore.case=T),
                                 "Sciences", DISCIPLINE)) %>%
        mutate(DISCIPLINE=ifelse(grepl('*Physic*',DISCIPLINE, ignore.case=T),
                                 "Sciences", DISCIPLINE)) %>%
        mutate(DISCIPLINE=ifelse(grepl('*Math*',DISCIPLINE, ignore.case=T),
                                 "Sciences", DISCIPLINE)) %>%
        mutate(DISCIPLINE=ifelse(grepl('*English*',DISCIPLINE, ignore.case=T),
                                 "HASS", DISCIPLINE)) %>%
        mutate(DISCIPLINE=ifelse(grepl('*Psychology*',DISCIPLINE, ignore.case=T),
                                 "HASS", DISCIPLINE)) %>%
        mutate(DISCIPLINE=ifelse(grepl('*Literature*',DISCIPLINE, ignore.case=T),
                                 "HASS", DISCIPLINE)) %>%
        mutate(DISCIPLINE=ifelse(grepl('*Sociology*',DISCIPLINE, ignore.case=T),
                                 "HASS", DISCIPLINE)) %>%
        mutate(DISCIPLINE=ifelse(grepl('*Media*',DISCIPLINE, ignore.case=T),
                                 "HASS", DISCIPLINE)) %>%
        mutate(DISCIPLINE=ifelse(grepl('*geog*',DISCIPLINE, ignore.case=T),
                                 "HASS", DISCIPLINE)) %>%
        mutate(DISCIPLINE=ifelse(grepl('*Music*',DISCIPLINE, ignore.case=T),
                                 "Audio-related", DISCIPLINE)) %>%
        mutate(DISCIPLINE=ifelse(grepl('*Sound*',DISCIPLINE, ignore.case=T),
                                 "Audio-related", DISCIPLINE)) %>%
        mutate(DISCIPLINE=ifelse(grepl('*Audio*',DISCIPLINE, ignore.case=T),
                                 "Audio-related", DISCIPLINE)) %>%
        mutate(DISCIPLINE=ifelse(grepl('*Acoustic*',DISCIPLINE, ignore.case=T),
                                 "Audio-related", DISCIPLINE)) %>%
        mutate(DISCIPLINE=ifelse(grepl('*engineering*',DISCIPLINE, ignore.case=T),
                                 "Engineering", DISCIPLINE)) %>%
        mutate(DISCIPLINE=ifelse(grepl('*tech*',DISCIPLINE, ignore.case=T),
                                 "Engineering", DISCIPLINE)) %>%
        mutate(DISCIPLINE=ifelse(!grepl('Sciences|Engineering|HASS|Audio-related',
                                       DISCIPLINE, ignore.case=F),
                                 "Others", DISCIPLINE))

```

### Formulation of criteria scores

Main Axes: APPR, UNDR, CLAR, ANTO, ORTH, NCON, IBAL
Derived Axes: APPR, UNDR, CLAR, CONN, IBAL

```{r formulation, message=FALSE}

#compute formulation across df of attributes on main axes
main.formulated<-list(eventdf,pleasdf,uneventdf,annoydf) %>%
        lapply(mainForm)

#compute formulation across df of attributes on derived axes                     
der.formulated<-list(vibrantdf,calmdf,monotdf,chaoticdf) %>%
        lapply(derForm)

eventformdf<-main.formulated[[1]]
pleasformdf<-main.formulated[[2]]
uneventformdf<-main.formulated[[3]]
annoyformdf<-main.formulated[[4]] %>%
        mutate(SPLITCANDIDATE=paste(COUNTRY,CANDIDATE))
vibrantformdf<-der.formulated[[1]] %>%
        mutate(SPLITCANDIDATE=paste(COUNTRY,CANDIDATE))
calmformdf<-der.formulated[[2]] %>%
        mutate(SPLITCANDIDATE=paste(COUNTRY,CANDIDATE))
monotformdf<-der.formulated[[3]] %>%
        mutate(SPLITCANDIDATE=paste(COUNTRY,CANDIDATE))
chaoticformdf<-der.formulated[[4]] %>%
        mutate(SPLITCANDIDATE=paste(COUNTRY,CANDIDATE))

```

## Demographic

The demographics of the 63 participants are summarised based on "length of stay outside SG/MY", "disiciplines", "malay language proficiency", and "english language proficiency".


```{r demoplots, message=FALSE}
#LoS
sgmydata_demo %>%
  group_by(LoSOut) %>%
  summarise(Singapore=sum(COUNTRY=="SG"),Malaysia=sum(COUNTRY=="MY")) %>%       
  kbl(caption = "Length of Stay Outside SG/MY") %>%
  kable_classic(full_width = F, html_font = "Cambria")

#Disicpline
sgmydata_demo %>%
  group_by(DISCIPLINE) %>%
  summarise(Singapore=sum(COUNTRY=="SG"),Malaysia=sum(COUNTRY=="MY")) %>%       
  kbl(caption = "Discipline") %>%
  kable_classic(full_width = F, html_font = "Cambria")

#Malay Language Proficiency
sgmydata_demo %>%
  group_by(Prof.zsm) %>%
  summarise(Singapore=sum(COUNTRY=="SG"),Malaysia=sum(COUNTRY=="MY")) %>%       
  kbl(caption = "Malay Language Proficiency") %>%
  kable_classic(full_width = F, html_font = "Cambria")

#English Language Proficiency
sgmydata_demo %>%
  group_by(Prof.zsm) %>%
  summarise(Singapore=sum(COUNTRY=="SG"),Malaysia=sum(COUNTRY=="MY")) %>%       
  kbl(caption = "English Language Proficiency") %>%
  kable_classic(full_width = F, html_font = "Cambria")

```

## Exploratory analysis

Compute criteria mean scores of all PAQ attributes across the combined, SG, and MY populations.

```{r explore,message=FALSE}

#Combined cases

#main axes
mainAxCOMBmean <- rbind(cbind(data.frame(PAQ="eventful"),mainAxSummary(eventformdf)),
      cbind(data.frame(PAQ="pleasant"),mainAxSummary(pleasformdf)),
      cbind(data.frame(PAQ="uneventful"),mainAxSummary(uneventformdf)),
      cbind(data.frame(PAQ="annoying"),mainAxSummary(annoyformdf))) %>%
        mutate(COUNTRY="Combined")

#derived axes
derAxCOMBmean <-rbind(cbind(data.frame(PAQ="vibrant"),derAxSummary(vibrantformdf)),
      cbind(data.frame(PAQ="calm"),derAxSummary(calmformdf)),
      cbind(data.frame(PAQ="monotonous"),derAxSummary(monotformdf)),
      cbind(data.frame(PAQ="chaotic"),derAxSummary(chaoticformdf))) %>%
        mutate(COUNTRY="Combined")

#Split cases (SG/MY)
#main axes
mainAxSPLITmean <- rbind(cbind(data.frame(PAQ="eventful"),mainAxSummarySPLIT(eventformdf)),
      cbind(data.frame(PAQ="pleasant"),mainAxSummarySPLIT(pleasformdf)),
      cbind(data.frame(PAQ="uneventful"),mainAxSummarySPLIT(uneventformdf)),
      cbind(data.frame(PAQ="annoying"),mainAxSummarySPLIT(annoyformdf)))

#derived axes
derAxSPLITmean <-rbind(cbind(data.frame(PAQ="vibrant"),derAxSummarySPLIT(vibrantformdf)),
      cbind(data.frame(PAQ="calm"),derAxSummarySPLIT(calmformdf)),
      cbind(data.frame(PAQ="monotonous"),derAxSummarySPLIT(monotformdf)),
      cbind(data.frame(PAQ="chaotic"),derAxSummarySPLIT(chaoticformdf)))

#pivot dataframe for visualisation
summaryMean<-rbind(rbind(mainAxCOMBmean,mainAxSPLITmean) %>% 
        pivot_longer(cols=!c(PAQ,CANDIDATE,COUNTRY),
                                      names_to="CRITERIA",
                                      values_to="mean"),
        rbind(derAxCOMBmean,derAxSPLITmean) %>% 
        pivot_longer(cols=!c(PAQ,CANDIDATE,COUNTRY),
                                      names_to="CRITERIA",
                                      values_to="mean")) %>%
        group_by(PAQ,CRITERIA,CANDIDATE) %>%
        summarise(Combined=mean[COUNTRY=="Combined"],
                  SG=mean[COUNTRY=="SG"],
                  MY=mean[COUNTRY=="MY"])

#Plot table of mean scores
summaryMean %>%
        kbl(caption = "Mean evaluation scores for the 
            PAQ attributes for all evaluation criteria across combined, 
            SG and MY populations") %>%
        kable_classic(full_width = F, html_font = "Cambria")


```

## Statistical analysis

### Main axes: One translation candidate

Due to differences in sample size, Kruskal-Wallis test was adopted to examine the statistical differences between SG and MY populations for PAQ attributes with only a single translation candidate, i.e. *eventful*, *uneventful*, and *pleasant*.

```{r kwtsolo, message=FALSE}
#KWT for solo translation candidates
kwtSolo<-rbind(kwTest(eventformdf,type = "main") %>% mutate(PAQ="eventful"),
               kwTest(uneventformdf,type = "main") %>% mutate(PAQ="uneventful"),
               kwTest(pleasformdf,type = "main") %>% mutate(PAQ="pleasant"))

#Display KWT results
kwtSolo %>%
        group_by(PAQ,CRITERION) %>%
        mutate(pvalue=as.numeric(pvalue),
               effect=as.numeric(effect)) %>%
        kbl(caption = "Kruskal-Wallis p-value and effect sizes for eventful, uneventful, and pleasant",
            digits = 3, booktabs=T) %>%
        kable_classic(full_width = F, html_font = "Cambria")

```

### Main axes: Mulitple translation candidate


